### 2020-03-14 ListaFiguras.tex
Pdf con fotos

\documentclass{article}

\usepackage[lofdepth,lotdepth]{subfig}

%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

\usepackage{pgf}
% \usepgfplotslibrary{external} 
% \tikzexternalize

%\usepackage{subfigure}
\usepackage{preview}
% Packages to write pseudo-algorithms %
\usepackage{algorithm}
\usepackage{algorithmic}

% Tikz
\usepackage{stanli}
\usepackage[ugly]{units}
% \usetikzlibrary{decorations}
% \usetikzlibrary{arrows}
\usetikzlibrary{plotmarks}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Discretization_MPM}
  }
  \caption{Discretization used in the material point method.}
  \label{fig:MPM-discretization}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{0.4\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{Classic material point method algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \centering
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Piecewise linear (Q4).]{
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/MPM_Shape_Fun_dy}
    \label{fig:MPM_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Local \textit{max-ent} $\gamma=17$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_17_3_Shape_Fun_dy}
    \label{fig:LME_17.3_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[Uniform GIMP (uGIMP).]{
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/GIMP_Shape_Fun_dy}
    \label{fig:GIMP_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subfloat[Local \textit{max-ent} $\gamma=10$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_10_0_Shape_Fun_dy}
    \label{fig:LME_10.0_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
  \subfloat[Local \textit{max-ent} $\gamma=7$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_7_0_Shape_Fun_dy}
    \label{fig:LME_7.0_Shape_Fun}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
  \subfloat[Local \textit{max-ent} $\gamma=5$.]{
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun_dx}
    \includegraphics[width=0.23\columnwidth]{Figures/LME_5_0_Shape_Fun_dy}
    \label{fig:LME_5.0_Shape_Fun}
  }
  \caption{Comparison of the local \textit{max-ent} shape function for
    different values of $\gamma = \beta h^2$, the piecewise linear shape function
    and the uniform GIMP shape function. The picture shows the nodal value
    of each shape function and its derivatives evaluated in a material
    point located in the center of the domain.}
  \label{fig:LME_MPM}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Dyka_bar.tex}}
  \caption{Geometrical description of the Dyka \cite{Dyka1995} bar.}
  \label{fig:Dyka_Bar}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_05}
  }
  \caption{Velocity evolution in the bar left side. This picture
    shows a comparison of both time integration algorithms,
    the Forward-Euler (FE) and the Predictor-Corrector explicit (PCE).}
  \label{fig:Velocity-Dyka-PCE-FE}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Stress_FE_vs_PCE_CFL_05}
  }
  \caption{Stress evolution in the bar left side. This picture
    shows a comparison of both time integration algorithms,
    the Forward-Euler (FE) and the Predictor-Corrector explicit (PCE).}
  \label{fig:Stress-Dyka-PCE-FE}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{    \includegraphics[width=\textwidth]{./Figures/Velocity_MPM_LME_gamma_comparative}
  }
  \caption{Velocity evolution at the point in the bar left side.}
  \label{fig:Velocity-Dyka-LME-gamma-comparative}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{    \includegraphics[width=\textwidth]{./Figures/Velocity_MPM_vs_OTM_Dyka}
  }
  \caption{Velocity evolution at the point in the bar left side.}
  \label{fig:Velocity-Dyka-MPM-vs-OTM}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{    \includegraphics[width=\textwidth]{./Figures/Velocity_Q4_vs_uGIMP_LME_Dyka}
  }
  \caption{Velocity evolution at the point in the bar left side.}
  \label{fig:Velocity-Dyka-Q4-uGIMP-LME}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \input{Figures/Block.tex}}
  \caption{Geometrical description of a soil block }.}
  \label{fig:block}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Block_CFL_01_Comparative}
  }
  \caption{Comparative of the vertical displacement evolution in a
    point located in the free surface employing different
    interpolation schemes and numerical techniques.} 
  \label{fig:vertical-displacement-block}
\end{figure}

\begin{figure}
  \centering
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[t = 0 seconds.]{
    \includegraphics[width=0.5\columnwidth]{Figures/Block_LME3_PCE_a_t0}
    \label{fig:Block-LME3-PCE-t0}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[t = 5 seconds.]{
    \includegraphics[width=0.5\columnwidth]{Figures/Block_LME3_PCE_b_t025}
    \label{fig:Block-LME3-PCE-t025}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \qquad
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \subfloat[t = 20 seconds]{
    \includegraphics[width=0.5\columnwidth]{Figures/Block_LME3_PCE_c_t1}
    \label{fig:Block-LME3-PCE-t1}
  }
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
  \caption{Vertical normal stress and position of material points
    during the loading process for a soft soil ($E = 5\ MPa$, $\rho_0
    = 6\cdot 10^3\ kg/m^3$). Numerical parameters considered for the
    simulation are : Local \textit{max-ent} shape function $\gamma =3$
    and explicit PC scheme with CFL 0.1.}
  \label{fig:Block-LME3}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\columnwidth]{Figures/1D_right_Velocity.png}
  }
  \caption[Velocities values in the right side of the Dyka
  bar]{Analytical solution for the velocity in the right side of the Dyka bar.}
  \label{fig:vel_analytics_dyka}
\end{figure}

\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\columnwidth]{Figures/1D_left_Stress.png}
  }
  \caption[Stress values in the last quarter side of the Dyka
  bar]{Analytical solution for the stress in the last quarter of the Dyka bar.}
  \label{fig:stress_analytics_dyka}
\end{figure}

\end{document}